Type Refinements for the Working Class

### 2016-01-01 paper.tex
introduce → induce
add nominal independence modality (#13)

An $\Sorts$-\emph{operator signature} is an $\Sorts$-arity-indexed family of copresheaves on
$\SCtx$, namely $\HypJ{\Of{\Operators_a}{\Sets^\SCtx}}{\IsArity{a}}$; this is
an intensional generalization of the state of affairs for standard universal
algebra, where operators are arity-indexed families of \emph{sets}. The point
of fibering over symbol contexts is to introduce an open-ended indexing
mechanism for operators; the functorial liftings of $\SCtx$-morphisms
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
equip operators $\Member{\vartheta}{\Operators_a(\Upsilon)}$ with a suitable
renaming action $\Member{\Rename{\rho}{\vartheta}}{\Operators_a(\Upsilon')}$.%
%
\footnote{
  In~\cite{sterling-morrison:2015}, the authors defined operator signatures as
  \emph{sheaves} on $\OpCat{(\SCtx)}$ under the atomic topology (alternatively,
  pullback-preserving presheaves), where all non-empty sieves cover; the sheaf
  condition induces the unique existence of \emph{strengthenings} for
  operators. However, because we do not use strengthening in the present
  development, we have omitted this part of the apparatus for the sake of
  simplicity.
}
%
We call the pair $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ an \emph{abt signature}.

@@

\subsubsection{The type of atoms}
\newcommand\SortAtom{\mathtt{atom}}
\newcommand\OpToken[1]{\mathtt{tok}[#1]}
\newcommand\OpMatch[1]{\mathtt{match}_{#1}}
\newcommand\Atoms{\mathbb{A}}

@@

\section{Related Work}

@@

>>>>

An $\Sorts$-\emph{operator signature} is an $\Sorts$-arity-indexed family of copresheaves on
$\SCtx$, namely $\HypJ{\Of{\Operators_a}{\Sets^\SCtx}}{\IsArity{a}}$; this is
an intensional generalization of the state of affairs for standard universal
algebra, where operators are arity-indexed families of \emph{sets}. The point
of fibering over symbol contexts is to induce an open-ended indexing
mechanism for operators; the functorial liftings of $\SCtx$-morphisms
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
equip operators $\Member{\vartheta}{\Operators_a(\Upsilon)}$ with a suitable
renaming action $\Member{\Rename{\rho}{\vartheta}}{\Operators_a(\Upsilon')}$.%
%
\footnote{
  In~\cite{sterling-morrison:2015}, the authors defined operator signatures as
  \emph{sheaves} on $\OpCat{(\SCtx)}$ under the atomic topology (alternatively,
  pullback-preserving presheaves), where all non-empty sieves cover; the sheaf
  condition induces the unique existence of \emph{strengthenings} for
  operators. However, because we do not use strengthening in the present
  development, we have omitted this part of the apparatus for the sake of
  simplicity.
}
%
We call the pair $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ an \emph{abt signature}.

@@

\subsubsection{The type of atoms}
\label{sec:atom-type}
\newcommand\SortAtom{\mathtt{atom}}
\newcommand\OpToken[1]{\mathtt{tok}[#1]}
\newcommand\OpMatch[1]{\mathtt{match}_{#1}}
\newcommand\Atoms{\mathbb{A}}

@@

\subsection{Refinement Modalities}

In Section~\ref{sec:atom-type} we introduced our first \emph{intensional} type,
that is, a type whose meaning varies over worlds $\Upsilon$. It is also
possible to introduce \emph{refinement modalities} which change the
meaning of a refinement in an intensional manner.

\subsubsection{Nominal Independence}

\newcommand\Indep[2]{#1\setminus #2}

Mark Bickford introduced in~\cite{bickford:2008} a type in Nuprl that expresses
the independence of a term $M$ of type $A$ from a symbol $a$, written
$M:A\parallel a$, which is used in order to express the notion that some agent
is not aware of a particular token/atom. In our setting, nominal independence
is best expressed through a comonadic refinement modality, $\Indep{\phi}{a}$, whose
inhabitants are the inhabitants of $\phi$ which do not make use of the symbol
$a$. To be precise, we intend to explain the following refinement formation
rule:
\[
  \infer={
    \Refines{\Upsilon,a:\sigma}{\Indep{\phi}{a}}{\tau}
  }{
    \Refines{\Upsilon,a:\sigma}{\phi}{\tau}
  }
\]

The definition of nominal independence is as follows, for any renaming
\begin{tikzcd}[cramped]
  \IMode{\Upsilon,a:\sigma}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd}%
:
\[
  \infer={
    \RMemEq{\Upsilon'}{M}{N}{\Indep{\phi}{\rho(a)}}
  }{
    \exists\Member{L}{\Exprs_\tau(\Upsilon'\setminus \rho(a))}.
    \begin{array}{l}
      \quad \RMemEq{\Upsilon'}{L}{M}{\phi}\\
      \quad \RMemEq{\Upsilon'}{L}{N}{\phi}
    \end{array}
  }
\]

In other words, a construction is independent of a symbol when it is equal to
one which was constructed without knowledge of that symbol.

\begin{thm}
  The refinement operator $\Indep{-}{a}$ is an idempotent comonad on refinements
  in the sense that the following rules are valid:

  \begin{gather*}
    \infer[\textit{extract}]{
      \RMemEq{\Upsilon}{M}{N}{\phi}
    }{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
    }
    \qquad
    \infer=[\textit{duplicate}]{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\Indep{\phi}{a}}{a}}
    }{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
    }
    \\[6pt]
    \infer[\textit{lift}]{
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
      \implies
      \RMemEq{\Upsilon}{M}{N}{\Indep{\psi}{a}}
    }{
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
      \RMemEq{\Upsilon}{M}{N}{\phi}
      \implies
      \RMemEq{\Upsilon}{M}{N}{\psi}
    }
  \end{gather*}
\end{thm}
\begin{proof}
  The first two inferences are obviously evident. To validate \textit{lift},
  suppose that $\RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}$; by inversion, we
  have an $\Member{L}{\Exprs_\tau(\Upsilon\setminus a)}$ such that
  $\RMemEq{\Upsilon}{L}{M}{\phi}$ and $\RMemEq{\Upsilon}{L}{N}{\phi}$.  Now, it
  suffices to exhibit an $\Member{L'}{\Exprs_\tau(\Upsilon\setminus a)}$ such
  that $\RMemEq{\Upsilon}{L'}{M}{\psi}$ and $\RMemEq{\Upsilon}{L'}{N}{\psi}$;
  but we may simply choose $\Match{L'}{L}$ and discharge the remaining subgoals
  by appealing to the premise of the rule.
\end{proof}

%% \subsubsection{Nominal Necessity}
%%
%% \subsubsection{Nominal Possibility}
%%
\section{Related Work}

@@

\section*{Acknowledgements}

The first author wishes to thank Robert Harper for the inspiration to pursue a
generalization of the Nuprl type theory from the refinements perspective. We
also thank Mark Bickford for clarifying the status of atoms and generative
phenomena in Nuprl; finally, we owe a debt of gratitude to Vincent Rahli and
Abhishek Anand for their tireless work on the \emph{Nuprl in Coq} project,
which was helpful in our efforts to understand Howe's computational equivalence
and its generalization to the nominal setting.
